Enabling the Use of GPU Virtualization in Cloud Environments

### 2015-10-08 paper.tex
comentado el discussion
varios

\keywords{Cloud Computing, GPU Virtualization, Resource Management}

\abstract{The use of accelerators in order to reduce the execution time of some 
application has become more popular during last years. These devices increment 
the computational power of a node thanks to its parallel architecture. %The popularity of the GPGPU computation 
This trend has led several cloud vendors such as Amazon Web Services or some 
middlewares such as OpenStack to add to their current facilities, or to allow, 
instances of virtual machines equipped with NVIDIA GPUs. 
Accomplishing this requirements needs that the guest hosts must be equipped with 
GPUs which could be barely utilized if a non GPU-enabled Virtual Machine is running in the host.
The solution presented in this work is based on GPU virtualization and shareability 
in order to reach an equilibrium between supply and demand of accelerators from the users. 
Hence, we propose to decouple from the nodes, real GPUs by using the virtualization technology called rCUDA. 
With this software configuration, these GPUs can be accessed from any VM getting rid 
of the necessity of having physical GPUs in the guest host. 
Moreover, our project studies the viability of these join by using a public cloud 
service configuration and develops a module for OpenStack middleware in order to 
add the needed support for the virtualized devices and the logic to manage them. 
The results show a feasible configuration which adds flexibility to the current and well-known 
cloud solutions. 
}

@@

\section{\uppercase{Background}}
\label{sec:background}
\subsection{The rCUDA Framework}
\label{sec:rcuda}

{rCUDA}~\cite{tonithesis,toniparco} is a middleware that enables transparent access
to any NVIDIA GPU device present in a cluster from all compute
nodes. The GPUs can be accessed and shared between nodes, and a single node can use all these graphic accelerators
as if they were local.
These features are focused in attaining higher accelerator utilization rates in the overall system while simultaneously reducing
resource, space, and energy~\cite{energy14} requirements.
rCUDA is structured following a client-server distributed
architecture: the client middleware is allocated in the same cluster node where the application demanding GPGPU
acceleration services is executed, providing a transparent replacement for the
native CUDA libraries. Furthermore, the server middleware is executed in the
cluster nodes from which the actual GPUs provide the requested GPGPU service.
To support a concurrent scenario where GPUs are shared between
processes\slash nodes, {rCUDA} manages separate device contexts for
each client application.

@@

An instance is a pre-configured virtual machine focused on an 
specific target. Inside the large list of instances offered by AWS, 
we can find: general pourpose (T2, M4 and M3); computer science optimized 
(C4 and C3); memory optimized (R3); storage optimized (I2 and D2) and 
GPU capable instances (G2). Each kind of instance has its own pourpose and its 
own price. Moreover, each instance has different number of CPUs and network 
interconnection which can be: low, medium, high or 10Gb.

@@

The IPERF tool has been executed between the instances showed in 
the Table~\ref{table:iperf} in order to know the real bandwidth.
Therefore, we can assure that the nomenclature ``High'' used in the 
network column correspond to a 1~Gb interconnection and the 10~Gb has 
a real bandwidth of 7.5~Gb/s.
Moreover, it looks like the instances equipped with a ``High'' interconnection 
are software-limited up to 1~Gb/s due to the theoretical and real bandwith 
match perfectly. The gap between real and theorical can be observed with 
the 10~Gb interconnection which reaches up to 7.5~Gb/s.

@@

\subsection{Testbed Scenarios}
In this section, the different client-server configurations 
are explained. All the scenarios are based in the maximum number of instance 
that a user can freely select without submiting a formal request.

@@

The first of them was {\tt MonteCarloMultiGPU} from NVIDIA SDK, launched with the default configuration ``scaling=weak'' which adjust the size of the problem depending on the number of accelerators.
The Figure~\ref{fig:mont-opt} depicts the options per second calculated by the application running on the scenarios of Figure~\ref{fig:aws} and on the instances with local GPUs. 
We have removed the results of the Scenario~\ref{subfig:aws2} for clarity ought to they are exactly the same that the obtained from Scenario~\ref{subfig:aws3} up to 8 GPUs. 
In other words, we could state that the results until the 8th GPU would correspond to the Scenario~\ref{subfig:aws2} and thereafter to the Scenario~\ref{subfig:aws3}.
Hence, in this particular case, rCUDA (using remote GPUs) is performing better than CUDA (with its local GPUs) due to rCUDA loads the libraries when its daemon starts.
However, even with rCUDA we can see differences in the results between booth scenarios. We determined that the Scenario~\ref{subfig:aws1} could increase the throughput because
unlike the other setups the GPUs are not sharing the PCI bus with other devices, as it happens in the 4-GPU instances.
% \begin{figure}[!t]
%   \centering
%   \includegraphics[width=\linewidth]{images/aws-mont2.png}
%   \caption{Montecarlo execution time}
%   \label{fig:mont-exe}
% \end{figure}
\begin{figure}[htb]
  \centering
  \includegraphics[width=\linewidth]{images/mont.pdf}
  \caption{NVIDIA SDK {\tt MonteCarloMultiGPU} execution using local and remote GPUs.}
  \label{fig:mont-opt}
\end{figure}

@@

\subsection{Experimental Results}
The tests were executed on a 5-node cluster where a machine
equipped with a Intel Xeon E7420 quadcore processor at
2.13 GHz, and 16 Gbytes of DDR2 RAM memory at 667
MHz, was guest host for the VMs. While the other 4 nodes
were auxiliary servers with a Tesla C1090 GPU each one. 
In the software stack the testbed system operates a
Centos 6.6; the cloud platform is deployed by OpenStack; and
the GPUs use CUDA 6.5. The GPU virtualization support is
based on rCUDA v5.0.

>>>>

\keywords{Cloud Computing, GPU Virtualization, AWS, Resource Management}

\abstract{The use of accelerators in order to reduce the execution time of some 
application has become more popular during last years. These devices increment 
the computational power of a node thanks to its parallel architecture.
This trend has led several cloud vendors such as Amazon Web Services or some 
middlewares such as OpenStack to add to their current facilities, or to allow, 
instances of virtual machines equipped with NVIDIA GPUs. 
Accomplishing this requirements needs that the guest hosts must be equipped with 
GPUs which could be barely utilized if a non GPU-enabled Virtual Machine is running in the host.
The solution presented in this work is based on GPU virtualization and shareability 
in order to reach an equilibrium between supply and demand of accelerators from the users. 
Hence, we propose to decouple from the nodes, real GPUs by using the virtualization technology called rCUDA. 
With this software configuration, these GPUs can be accessed from any VM getting rid 
of the necessity of having physical GPUs in the guest host. 
Moreover, our project studies the viability of these join by using a public cloud 
service configuration and develops a module for OpenStack middleware in order to 
add the needed support for the virtualized devices and the logic to manage them. 
The results show a feasible configuration which adds flexibility to the current and well-known 
cloud solutions. 
}

@@

\section{\uppercase{Background}}
\label{sec:background}
\subsection{The rCUDA Framework}
\label{sec:rcuda}
{rCUDA}~\cite{tonithesis,toniparco} is a middleware that enables transparent access
to any NVIDIA GPU device present in a cluster from all compute
nodes. The GPUs can be accessed and shared between nodes, and a single node can use all these graphic accelerators
as if they were local.
These features are focused in attaining higher accelerator utilization rates in the overall system while simultaneously reducing
resource, space, and energy~\cite{energy14} requirements.
rCUDA is structured following a client-server distributed
architecture: the client middleware is allocated in the same cluster node where the application demanding GPGPU
acceleration services is executed, providing a transparent replacement for the
native CUDA libraries. Furthermore, the server middleware is executed in the
cluster nodes from which the actual GPUs provide the requested GPGPU service.
To support a concurrent scenario where GPUs are shared between
processes\slash nodes, {rCUDA} manages separate device contexts for
each client application.

@@

An instance is a pre-configured virtual machine focused on an 
specific target. Inside the large list of instances offered by AWS, 
we can find: general pourpose (T2, M4 and M3); computer science optimized 
(C4 and C3); memory optimized (R3); storage optimized (I2 and D2) and 
GPU capable instances. Each kind of instance has its own pourpose and its 
own price. Moreover, each instance has different number of CPUs and network 
interconnection which can be: low, medium, high or 10Gb.

@@

The IPERF tool has been executed between the instances showed in 
the Table~\ref{table:iperf} in order to know the real bandwidth.
Therefore, we can assure that the nomenclature ``High'' used in the 
network column correspond to a 1Gb interconnection and the 10Gb has 
a real bandwidth of 7.5Gb/s.
Moreover, it looks like the instances equipped with a ``High'' interconnection 
are software-limited up to 1 Gb/s due to the theoretical and real bandwith 
match perfectly. The gap between real and theorical can be observed with 
the 10Gb interconnection which reaches up to 7.5 Gb/s.

@@

\subsection{Testbed Scenarios}
In this section, the different client-server configurations 
are explained. All the scenarios are based in the maximum number of instances 
that a user can freely select without submitting a formal request. 
In particular, can be launched 5 of ``g2.2xlarge'' and 2 of ``g2.8xlarge''.

@@

The first of them was {\tt MonteCarloMultiGPU} from NVIDIA SDK, an application purely computational. 
It was launched with the default configuration ``scaling=weak'' which adjust the size of the problem depending on the number of accelerators.
The Figure~\ref{fig:mont-opt} depicts the options per second calculated by the application running on the scenarios of Figure~\ref{fig:aws} and on the instances with local GPUs. 
We have removed the results of the Scenario~\ref{subfig:aws2} for clarity ought to they are exactly the same that the obtained from Scenario~\ref{subfig:aws3} up to 8 GPUs. 
In other words, we could state that the results until the 8th GPU would correspond to the Scenario~\ref{subfig:aws2} and thereafter to the Scenario~\ref{subfig:aws3}.
Hence, in this particular case, rCUDA (using remote GPUs) is performing better than CUDA (with its local GPUs) due to rCUDA loads the libraries when its daemon starts.
However, even with rCUDA we can see differences in the results between booth scenarios. We determined that the Scenario~\ref{subfig:aws1} could increase the throughput because
unlike the other setups the GPUs are not sharing the PCI bus with other devices, as it happens in the 4-GPU instances.
% \begin{figure}[!t]
%   \centering
%   \includegraphics[width=\linewidth]{images/aws-mont2.png}
%   \caption{Montecarlo execution time}
%   \label{fig:mont-exe}
% \end{figure}
\begin{figure}[htb]
  \centering
  \includegraphics[width=\linewidth]{images/mont.pdf}
  \caption{NVIDIA SDK {\tt MonteCarloMultiGPU} execution using local and remote GPUs.}
  \label{fig:mont-opt}
\end{figure}

@@

\subsection{Experimental Results}
The tests were executed on a 5-node cluster where a machine
equipped with a Intel Xeon E7420 quadcore processor at
2.13 GHz, and 16 Gbytes of DDR2 RAM memory at 667
MHz, was guest host for the VMs. While the other 4 nodes
were auxiliary servers with a Tesla C1060 GPU each one. 
In the software stack the testbed system operates a
Centos 6.6; the cloud platform is deployed by OpenStack; and
the GPUs use CUDA 6.5. The GPU virtualization support is
based on rCUDA v5.0.
